---
title: "hw"
author: "Blaire(Moyuan) Li"
date: "2/6/2017"
output: html_document
---
Code up the generative model (i.e. generate data from) for a Dirichlet process mixture of Gaussians (preferably in R, Matlab, or Python), using a Normal-Inverse Wishart prior on the Gaussian parameters.

What happens as you vary the DP concentration parameter (alpha)? What happens as you vary the Normal-Inverse Wishart hyperparameters? Along with your code, submit plots of your generated data to back up your answers.

```{r}
library(ggplot2)
library(NPflow)
set.seed(100)
rDPGM = function(n,alpha,hyperG0){
  theta <- list() # the list that stores the gaussian parameters for each cluster
  C <- numeric(n) # specifies which cluster that each point belongs to
  z <- matrix(NA, ncol=n, nrow=d) # stores the generated points
  
  for (i in 1:n){
    p0 <- alpha/(i-1+alpha) #the probability that sits down at a new table
    u <- runif(n=1, min = 0, max = 1)
    if (u<p0){
      # choose a new cluster
      C[i] <- max(C)+1 
      theta[[C[i]]] <- rNNiW(hyperG0$NNiW, diagVar=FALSE)
      theta[[C[i]]][["nu"]] <- 1 + rgamma(n=1, shape = 2, rate=1/10)
      
    }else{
      # choose an exisitng cluster
      u1 = u-p0
      weights <- summary(factor(C))[-1] # the N_k for the each cluster except the first cluster
      C[i] <- which(rmultinom(n=1, size=1, prob=weights)==1)
    }
    w <- rgamma(n=1, shape=theta[[C[i]]]$nu/2, rate=theta[[C[i]]]$nu/2)
    ltnz <- rtruncnorm(n=1, a=0, sd=1/sqrt(w))
    eps <- matrix(rnorm(d), ncol=d)%*%chol(theta[[C[i]]]$S/w)
    z[1,i] <- theta[[C[i]]]$xi[1]+theta[[C[i]]]$psi[1]*ltnz+eps[,1]
    z[2,i] <- theta[[C[i]]]$xi[2]+theta[[C[i]]]$psi[2]*ltnz+eps[,2] 
  }
  
  sim=cbind.data.frame("D1"=z[1,],"D2"=z[2,],"Cluster"=as.factor(C))
  return(sim)
}



#plots with varying alpha 
d=2
hyperG0 <- list()
hyperG0[["NNiW"]] <- list()
hyperG0[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG0[["NNiW"]][["b_psi"]] <- rep(0,d)
hyperG0[["NNiW"]][["D_xi"]] <- 100
hyperG0[["NNiW"]][["D_psi"]] <- 8
hyperG0[["NNiW"]][["nu"]] <- d+1
hyperG0[["NNiW"]][["lambda"]] <- diag(c(1,1))
n= 1000   #number of observations
sim1 = rDPGM(1000, 1,hyperG0)
sim2 = rDPGM(1000, 5,hyperG0)
sim3 = rDPGM(1000, 15,hyperG0)
sim4 = rDPGM(1000, 30,hyperG0)
p1 = (ggplot(data=sim1,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p1
p2 = (ggplot(data=sim2,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p2
p3 = (ggplot(data=sim3,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p3
p4 = (ggplot(data=sim4,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p4

```


### The points clump together in fewer clusters for low values of alpha, but become more dispersed as alpha increases.

```{r}
d=2
hyperG0 <- list()
hyperG0[["NNiW"]] <- list()
hyperG0[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG0[["NNiW"]][["b_psi"]] <- rep(0,d)
hyperG0[["NNiW"]][["D_xi"]] <- 100
hyperG0[["NNiW"]][["D_psi"]] <- 3
hyperG0[["NNiW"]][["nu"]] <- d+1
hyperG0[["NNiW"]][["lambda"]] <- diag(c(1,1))

hyperG1 <- list()
hyperG1[["NNiW"]] <- list()
hyperG1[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG1[["NNiW"]][["b_psi"]] <- rep(0,d)
hyperG1[["NNiW"]][["D_xi"]] <- 100
hyperG1[["NNiW"]][["D_psi"]] <- 16
hyperG1[["NNiW"]][["nu"]] <- d+1
hyperG1[["NNiW"]][["lambda"]] <- diag(c(1,1))

hyperG2 <- list()
hyperG2[["NNiW"]] <- list()
hyperG2[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG2[["NNiW"]][["b_psi"]] <- rep(0,d)
hyperG2[["NNiW"]][["D_xi"]] <- 100
hyperG2[["NNiW"]][["D_psi"]] <- 40
hyperG2[["NNiW"]][["nu"]] <- d+1
hyperG2[["NNiW"]][["lambda"]] <- diag(c(1,1))


hyperG3 <- list()
hyperG3[["NNiW"]] <- list()
hyperG3[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG3[["NNiW"]][["b_psi"]] <- rep(0,d)
hyperG3[["NNiW"]][["D_xi"]] <- 100
hyperG3[["NNiW"]][["D_psi"]] <- 1000
hyperG3[["NNiW"]][["nu"]] <- d+1
hyperG3[["NNiW"]][["lambda"]] <- diag(c(1,1))




set.seed(200)
n= 1000   #number of observations
sim1 = rDPGM(1000, 1,hyperG0)
sim2 = rDPGM(1000, 1,hyperG1)
sim3 = rDPGM(1000, 1,hyperG2)
sim4 = rDPGM(1000, 1,hyperG3)
p1 = (ggplot(data=sim1,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p1
p2 = (ggplot(data=sim2,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p2
p3 = (ggplot(data=sim3,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p3
p4 = (ggplot(data=sim4,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p4

```

####Increasing D_psi seems to makes shape of the cluster longer and thinner.

```{r}
d=2
hyperG0 <- list()
hyperG0[["NNiW"]] <- list()
hyperG0[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG0[["NNiW"]][["b_psi"]] <- rep(0,d)
hyperG0[["NNiW"]][["D_xi"]] <- 100
hyperG0[["NNiW"]][["D_psi"]] <- 16
hyperG0[["NNiW"]][["nu"]] <- d+1
hyperG0[["NNiW"]][["lambda"]] <- diag(c(1,1))

hyperG1 <- list()
hyperG1[["NNiW"]] <- list()
hyperG1[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG1[["NNiW"]][["b_psi"]] <- rep(3,d)
hyperG1[["NNiW"]][["D_xi"]] <- 100
hyperG1[["NNiW"]][["D_psi"]] <- 16
hyperG1[["NNiW"]][["nu"]] <- d+1
hyperG1[["NNiW"]][["lambda"]] <- diag(c(1,1))

hyperG2 <- list()
hyperG2[["NNiW"]] <- list()
hyperG2[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG2[["NNiW"]][["b_psi"]] <- rep(10,d)
hyperG2[["NNiW"]][["D_xi"]] <- 100
hyperG2[["NNiW"]][["D_psi"]] <- 16
hyperG2[["NNiW"]][["nu"]] <- d+1
hyperG2[["NNiW"]][["lambda"]] <- diag(c(1,1))


hyperG3 <- list()
hyperG3[["NNiW"]] <- list()
hyperG3[["NNiW"]][["b_xi"]] <- rep(0,d)
hyperG3[["NNiW"]][["b_psi"]] <- rep(100,d)
hyperG3[["NNiW"]][["D_xi"]] <- 100
hyperG3[["NNiW"]][["D_psi"]] <- 16
hyperG3[["NNiW"]][["nu"]] <- d+1
hyperG3[["NNiW"]][["lambda"]] <- diag(c(1,1))




set.seed(200)
n= 1000   #number of observations
sim1 = rDPGM(1000, 1,hyperG0)
sim2 = rDPGM(1000, 1,hyperG1)
sim3 = rDPGM(1000, 1,hyperG2)
sim4 = rDPGM(1000, 1,hyperG3)
p1 = (ggplot(data=sim1,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p1
p2 = (ggplot(data=sim2,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p2
p3 = (ggplot(data=sim3,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p3
p4 = (ggplot(data=sim4,aes(x=D1, y=D2))
      + geom_point(aes(colour=Cluster), alpha=0.6)
      + theme_bw()
        )
p4

```

####Increasing the b_psi seems to make the cluster more parallel
 reference for some codes:
 https://github.com/borishejblum/NPflow/blob/master/R/rCRP.R
